Supporting Information

### 2015-10-23 ms.Rmd
Moved MADS figure to SI. Needs cleaning.

# Results and Discussion

@@

| **Shorten MADS section and insert here. MADS DEG figure goes to SI?**
| **GRAS TFs (SHR1/SCR)? I think I will skip this because they are not highly expressed.**

@@

| **[Re]move following MADS section and insert section on Homeobox. Can link via WUS (HB, target of cytokinin, e.g. Han mini-review) and/or OSH1 (HB, downregulated in LOG mutant, e.g. Kurakawa 2007)**

@@

## Comparative analysis of MADS gene expression in eudicots and monocots

@@

* Something about MADS genes involved earlier in development?
* Expression of MADS genes that interact with *G1L5*? (Maybe this is too much detail?)

@@

@@

```{r f_madsTree}
print(f_madsTree)
```

**`r I(pasteLabel("Figure ", figCount, "f_madsTree"))`**. Differential expression of putatative MIKC proteins in *A. thaliana*, *Solanum lycopersicum* and *Oryza sativa* ssp. *japonica*. The tree is a neighbour-joining tree estimated from genetic distance of amino acid sequences. Named proteins starting with AT_ or OS_ are from Arabidopsis or rice respectively, with synonyms in parentheses, and other proteins are labelled with their unique identifiers. The colour of the label represents the Log~2~-fold change (L~2~FC) in expression of the gene encoding that protein between undifferentiated and differentiated meristems. Where there is no coloured label, the adjusted *p*-value associated with the difference in expression was greater than 0.1, or the gene for the protein in question was excluded from the differential expression analysis by `DESeq2`, which may occur because of low read counts, an extreme count outlier or the gene failing independent filtering [@Love:2014do]. For rice, L~2~FC was estimated from the RM and SM samples reported here. For *A. thaliana*, L~2~FC was estimated using the inflorescence meristem and floret meristem samples reported by @Mantegazza:2014di. The sympodial inflorescence meristem and floret meristem samples generated by @Park:2012dg were used to estimate L~2~FC values for tomato genes. Clades are labelled following the nomenclature presented by @Arora:2007fa.

# `sessionInfo` (*hide*)

>>>>

# Results

@@

# Discussion

## Changes between apical and axillary meristems

## Gradual changes in expression during axillary meristem transitions

@@

| **Shorten MADS section and insert here. MADS DEG figure goes to SI?**

@@

@@

@@

## Dynamic expression of homeodomain genes

Can link via WUS (HB, target of cytokinin, e.g. Han mini-review) and/or OSH1 (HB, downregulated in LOG mutant, e.g. Kurakawa 2007).

@@

----
# `sessionInfo` (*hide*)

### 2015-10-23 si.Rmd
Moved MADS figure to SI. Needs cleaning.

---
title: "Supporting Information"
csl: csl/the-plant-journal.csl
header-includes:
  \usepackage{graphicx}
  \usepackage{caption}
  \captionsetup[figure]{labelfont=bf}
  \renewcommand{\thefigure}{S\arabic{figure}}
output:
  rmarkdown::tufte_handout:
    dev: "CairoPDF"
  pdf_document:
    fig_caption: yes
    number_sections: yes
    dev: "CairoPDF"
  BiocStyle::html_document:
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    number_sections: yes
    dev: "CairoPDF"
  word_document:
    fig_caption: yes
bibliography: bib/papersLib.bibtex
---

@@

knitr::opts_chunk$set(fig.path='figures/',
               cache.path='cache/',
#               fig.align = 'center', 
               dev = "cairo_pdf",
               dpi=300,
               fig.show='hold',
               fig.height = 3.150,
               fig.width = 3.150,
               out.width = 3.150,
               out.height = 3.150,
#               cache=FALSE,
               echo=FALSE,
               results="hide",
               message=FALSE,
               warning=FALSE)

pander::panderOptions("table.caption.prefix", "")
source('src/R/figures.R')

@@

``` {r st_libStats, results = "asis", fig.cap = I(pasteLabel("**Table S", s_tableCount, "st_libStats"))}
pander::set.caption(paste0(
  I(pasteLabel("**Table S", s_tableCount, "st_libStats")),
  "**: Library and sequencing statistics."))
pander::pandoc.table(st_libStats, split.tables = Inf, missing = "", round = 1, emphasize.strong.cols = 1,
                     justify = c("right", rep("center", length(names(st_libStats)) - 1)))
```

----

```{r st_reviewInSitu, results = "asis"}
pander::set.caption(paste0(
  I(pasteLabel("**Table S", s_tableCount, "st_reviewInSitu")),
  "**: Comparison of the dataset with published *in situ* expression profiles. '×' indicates that the gene was detected by both methods, '+' indicates that the gene was detected by *in situ* hybridisation but not by RNA sequencing and '·' indicates that the gene was detected by RNA sequencing but expression was not reported by *in situ*. The last category is not intended to be informative, because it relies on negative or unreported results. A blank cell indicates that the gene was not detected by either method."))
pander::pandoc.table(st_reviewInSitu, split.tables = Inf,
                     justify = c("right", rep("centre", length(names(st_reviewInSitu)) - 2), "left"),
                     emphasize.cols = c(1,2), missing = "")
```

----

```{r sf_pca, fig.cap = paste0("Principal components analysis of transformed read counts for each library. The two libraries made from degraded RNA did not cluster with other libraries (see ", I(pasteLabel("Table S", s_tableCount, "st_libStats")), "), suggesting that the use of degraded RNA affected quantification of expression. For this reason these two libraries were not included in subsequent analyses.")}
print(sf_pca)
```

**`r I(pasteLabel("Figure S", s_figCount, "sf_pca"))`**: Principal components analysis of transformed read counts for each library. The two libraries made from degraded RNA did not cluster with other libraries (see `r I(pasteLabel("Table S", s_tableCount, "st_libStats"))`), suggesting that the use of degraded RNA affected quantification of expression. For this reason these two libraries were not included in subsequent analyses.

----

```{r sf_isGenesTpm, fig.height = 8.661, fig.width = 6.614}
print(sf_isGenesTpm)
```

**`r I(pasteLabel("Figure S", s_figCount, "sf_isGenesTpm"))`**: Quantitative expression values for the genes that were compared to published *in situ* results. The expression value in transcripts per million is shown on the *x*-axis. Points are coloured by replicate, and plotted as a triangle if the gene was above the expression cutoff or a circle otherwise. Although four replicates were produced from each of the RM and PBM, only three replicates for each stage are included in the analysis (see Results).

----

```{r sf_mfuzzCentroids}
print(sf_mfuzzCentroids)
```

**`r I(pasteLabel("Figure S", s_figCount, "sf_mfuzzCentroids"))`**: Minimum centroid distance for 2–`r maxClust` cluster cores (*c*). Eight cluster cores were used because the rate of change of the minimum centroid distance over *c* decreases after *c* = 8.

----

``` {r sf_mfuzzPca}
print(sf_mfuzzPca)
```

**`r I(pasteLabel("Figure S", s_figCount, "sf_mfuzzPca"))`**: Classical multidimensional scaling (MDS) of Euclidean distances between transformed read counts for clustered genes. Genes are coloured by the cluster to which they were assigned. The separation of genes by cluster on the MDS plot supports the use of eight clusters.

----

>>>>

---
title: "Supporting Information"
csl: csl/the-plant-journal.csl
header-includes:
  \usepackage{graphicx}
  \usepackage{caption}
  \usepackage{fixltx2e}
  \usepackage{lscape}
  \usepackage{rotating}
  \captionsetup[figure]{labelfont=bf}
  \renewcommand{\thefigure}{S\arabic{figure}}
  \renewcommand{\tablename}{ \hfill \newline Table}
  \renewcommand{\thetable}{S\arabic{table}}
documentclass: tufte-handout
classoption: a4paper
output:
  rmarkdown::tufte_handout:
    dev: CairoPDF
    keep_tex: yes
  BiocStyle::html_document:
    dev: CairoPDF
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    number_sections: yes
    toc: yes
  pdf_document:
    fig_caption: yes
    number_sections: yes
    dev: "CairoPDF"
  word_document:
    fig_caption: yes
bibliography: bib/papersLib.bibtex
---

@@

knitr::opts_chunk$set(fig.path='figures/',
               cache.path='cache/',
               fig.align = 'center', 
#               dev=c("png", "cairo_pdf"),
               dpi=300,
               fig.show='hold',
               fig.height = 4.213,
               fig.width = 4.213,
               crop = FALSE,
#               cache=FALSE,
               echo=FALSE,
               results="hide",
               message=FALSE,
               warning=FALSE)

#pander::panderOptions("table.caption.prefix", "")
source('src/R/figures.R')

@@

``` {r st_libStats, results = "hide", fig.cap = I(pasteLabel("**Table S", s_tableCount, "st_libStats"))}

pander::set.caption(paste0(
  I(pasteLabel("**Table S", s_tableCount, "st_libStats")),
  "**: Library and sequencing statistics."))
pander::pandoc.table(st_libStats,
                     missing = "",
                     round = 1,
                     emphasize.strong.cols = 1,
                     split.tables = Inf,
                     justify = c("right", rep("center", length(names(st_libStats)) - 1)))

\hfill\newline
\hfill\newline
\begin{longtable}[c]{@{}cc@{}}
\caption{\protect\label{tab:st_libstats}Library and sequencing statistics.}
\end{longtable}

``` {r st_libstats2, results = "asis"}

st_libStats_x <- as.data.frame(st_libStats)
rownames(st_libStats_x) <- st_libStats_x$Library
st_libStats_x$Library <- NULL
st_libStats_x <- xtable::xtable(st_libStats_x)
xtable::digits(st_libStats_x) <- c(0, rep(1, ncol(st_libStats_x) - 1), 0)
#xtable::caption(st_libStats_x) <- "Library and sequencing statistics."

st_libStats_x_print <- print(st_libStats_x,
      caption.placement = "top",
      comment = FALSE,
      rotate.colnames = TRUE,
      floating = TRUE,
      floating.environment = "table*",
      print.results = FALSE)

cat(st_libStats_x_print)

#cat(sub("\\begin{table*}[ht]\n", "\\begin{table*}[ht]\n\\caption{Library and sequencing statistics.} \n", st_libStats_x_print, fixed = TRUE))

\pagebreak

```{r st_reviewInSitu, results = "hide"}
pander::set.caption(paste0(
  I(pasteLabel("**Table S", s_tableCount, "st_reviewInSitu")),
  "**: Comparison of the dataset with published *in situ* expression profiles. '×' indicates that the gene was detected by both methods, '+' indicates that the gene was detected by *in situ* hybridisation but not by RNA sequencing and '·' indicates that the gene was detected by RNA sequencing but expression was not reported by *in situ*. The last category is not intended to be informative, because it relies on negative or unreported results. A blank cell indicates that the gene was not detected by either method."))
pander::pandoc.table(st_reviewInSitu, split.tables = Inf,
                     justify = c("right", rep("centre", length(names(st_reviewInSitu)) - 2), "left"),
                     emphasize.cols = c(1,2), missing = "")
```

```{r sf_pca, fig.cap = paste0("Principal components analysis of transformed read counts for each library. The two libraries made from degraded RNA did not cluster with other libraries (see Table \\ref{tab:st_libstats}), suggesting that the use of degraded RNA affected quantification of expression. For this reason these two libraries were not included in subsequent analyses.")}
print(sf_pca)
```

```{r sf_isGenesTpm, fig.height = 9.535, fig.cap = paste0("\\protect\\label{fig:sf_isGenesTpm}","Comparison of the LMD RNA-sequencing dataset with published \\emph{in situ} results. Expression value in transcripts per million (TPM) is shown on the \\emph{y}-axis, and points are plotted as a triangle if the expression value was above the cutoff for that library or a circle if below. Grey shading indicates a meristem type for which expression was detected by \\emph{in situ} hybridisation. See Table \\ref{tab:refFrame_x} for references for the \\emph{in situ} expression patterns. [\\textbf{As this is now detable-ized include a separate supplemantary table of references}] Although four replicates were produced from each of the RM and PBM, only three replicates for each stage are included in the analysis (see Results).")}

print(sf_isGenesTpm)

\clearpage

```{r insituRefs, results = "asis"}

# refFrame_x <- print(xtable::xtable(refFrame,
#                                    caption = "References for Figure \\ref{fig:sf_isGenesTpm}",
#                                    label = "tab:refFrame_x"),
#       type = "latex",
#       floating = TRUE,
#       floating.environment = "table",
#       rotate.colnames = FALSE,
#       print.results = FALSE)

setnames(refTable, c("msuId", "plotLabel"), c("MSU name", "CSGNL symbol"))
pander::set.caption("\\protect\\label{tab:refFrame_x}References for Figure \\ref{fig:sf_isGenesTpm}.")
pander::pandoc.table(refTable, emphasize.cols = c(1,2))
```

\clearpage

```{r sf_mfuzzCentroids, fig.cap = paste0("Minimum centroid distance for 2–", maxClust, " cluster cores (\\emph{c}). Eight cluster cores were indicated because the rate of change of the minimum centroid distance over \\emph{c} decreases after \\emph{c} = 8."), fig.scap = "Something", fig.scap = c("One plot", "Another plot")}

print(sf_mfuzzCentroids)

``` {r sf_mfuzzPca, fig.cap = paste0("Classical multidimensional scaling (MDS) of Euclidean distances between transformed read counts for clustered genes. Genes are coloured by the cluster to which they were assigned. The separation of genes by cluster on the MDS plot supports the use of eight clusters.")}
print(sf_mfuzzPca)
```

```{r sf_madsTree, fig.height = 9.535, fig.cap = paste0("Differential expression of putatative MIKC proteins in \\emph{A. thaliana}, \\emph{Solanum lycopersicum} and \\emph{Oryza sativa} ssp. \\emph{japonica}. The tree is a neighbour-joining tree estimated from genetic distance of amino acid sequences. Named proteins starting with AT\\_ or OS\\_ are from Arabidopsis or rice respectively, with synonyms in parentheses, and other proteins are labelled with their unique identifiers.", "\\newline ", "The colour of the label represents the Log\\textsubscript{2}-fold change (L\\textsubscript{2}FC) in expression of the gene encoding that protein between undifferentiated and differentiated meristems. Where there is no coloured label, the adjusted \\emph{p}-value associated with the difference in expression was greater than 0.1, or the gene for the protein in question was excluded from the differential expression analysis by `DESeq2`, which may occur because of low read counts, an extreme count outlier or the gene failing independent filtering \\citep{@Love:2014do}.", "\\newline ", "For rice, L\\textsubscript{2}FC was estimated from the RM and SM samples reported here. For \\emph{A. thaliana}, L\\textsubscript{2}FC was estimated using the inflorescence meristem and floret meristem samples reported by \\citep{Mantegazza:2014di}. The sympodial inflorescence meristem and floret meristem samples generated by \\citep{Park:2012dg} were used to estimate L\\textsubscript{2}FC values for tomato genes. Clades are labelled following the nomenclature presented by \\citep{Arora:2007fa}.")}

print(sf_madsTree)
```

\clearpage
